---
output:
  html_document: default
  pdf_document: default
---

|                                    |
|------------------------------------|
| Title: "Ustar threshold detection and MDS gapfilling" |
| Author: "Fabio Turco"              |
| Date: "2024-04-24"                 |
| Output:                            |
| pdf_document: default              |
| html_document: default             |

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

**Script to detect seasonal Ustar thresholds from level-3.2 fluxes.\
**In addition, the script is conducting MDS gap-filling of NEE, LE and H.\
**Finally, night-time partitioning is conducted for 16, 50, and 84 NEE CUTs.

HIGHEST QUALITY ONLY (QCF=0) FLUXES (\_L3.1_L3.2_QCF0), quality-controlled with Level-2 and Level-3.2 flags, including Level-3.1 storage correction are used for the ustar detection.

## LIBRARIES
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(REddyProc)
library(caTools)
library(data.table)
library(viridis)
```

Define the run ID
```{r}
run_id <- format(Sys.time(), "%Y%m%d",tz="GMT")
print(run_id)
```

## SOURCE FILE
```{r}
file_fluxes_meteo <- "../40_FLUX_PROCESSING_CHAIN/46.2_FluxProcessingChain_L3.2_subset-forREddyProc.csv"
```

## DATA FROM FILE

-   Timestamp is in Etc/GMT-1
-   Need to provide an end timestamp because EddyProc is shifting to the middle internally by subtracting 15 minutes

```{r}
filedata <- read_csv(file_fluxes_meteo, show_col_types = FALSE)
# TIMESTAMP_END holds CET clock times, which readr parses as UTC: relabel as CET without shifting the clock
# (as.POSIXct(<POSIXct>, tz = ...) shifts the clock time in R >= 4.3)
filedata$TIMESTAMP_END <- lubridate::force_tz(as.POSIXct(filedata$TIMESTAMP_END, tz = "UTC"), tzone = "Etc/GMT-1")
```
## TIME INTERVAL

Dataset period: 17 Oct 2023 00:00 to 23 Aug 2024 00:00 (TIMESTAMP_END, CET). Must match `PERIOD_START` / `PERIOD_END` in `src/config.py`.

```{r}
filedata <- filedata %>% filter(TIMESTAMP_END > as.POSIXct('2023-10-17 00:00:00', tz = 'Etc/GMT-1')
                                & TIMESTAMP_END <= as.POSIXct('2024-08-23 00:00:00', tz = 'Etc/GMT-1'))

summary(filedata[c('TIMESTAMP_END', 'NEE_L3.1_L3.2_QCF', 'sw_in', 'ta', 'vpd')], 3)
```
## DATA COLUMNS

**Using QCF fluxes**
```{r}
EddyData.F <- filedata[FALSE]

EddyData.F$TIMESTAMP <- filedata$TIMESTAMP_END

EddyData.F$NEE <- filedata$NEE_L3.1_L3.2_QCF

EddyData.F$LE <- filedata$LE_L3.1_L3.2_QCF

EddyData.F$H <- filedata$H_L3.1_L3.2_QCF

EddyData.F$Ustar <- filedata$USTAR

EddyData.F$Rg <- filedata$sw_in

EddyData.F$Tair <- filedata$ta

EddyData.F$VPD <- filedata$vpd

```

## EDDYPROC CONFIGURATION

### Initialize R5 reference class
```{r}
EddyProc.C <-sEddyProc$new('CH-TAN',  EddyData.F,
                           c('NEE', 'Rg','Tair', 'VPD', 'LE', 'H', 'Ustar'), ColPOSIXTime = "TIMESTAMP")   
EddyProc.C$sSetLocationInfo(LatDeg = 47.480620, LongDeg = 8.911868, TimeZoneHour = 1)   
head(EddyProc.C$sDATA)
```

### Define seasons

Seasons defined based on the management events of the winter wheat season 2023-24:

-   on 290 2023 start of dataset (manure application; wheat sowing on DOY 291)
-   on 122 2024 wheat height >  30 cm
-   on 208 2024 wheat harvested (bare soil until the end of the dataset on 22 Aug 2024)

```{r}
y <- c(2023, 2024, 2024)
s <- c(290, 122, 208)
seasonStarts <- as.data.frame(cbind(s,y))
seasonFactor <- usCreateSeasonFactorYdayYear(EddyData.F$TIMESTAMP, starts=seasonStarts)
seasonStartsDate <-fConvertTimeToPosix(data.frame(Year=seasonStarts[, 2],
                                                  DoY=seasonStarts[, 1],
                                                  Hour=12),
                                       'YDH', Year = "Year", Day = "DoY", Hour = "Hour")
print(seasonStartsDate)
```

## SEASONAL USTAR DETECTION
```{r}
EddyProc.C$sEstimateUstarScenarios(

  seasonFactor = seasonFactor,
  nSample = 200L, # nsample: how many bootstrap samples you want for your data
  probs = c(0.16, 0.5, 0.84)
  )

# display ustar scenarios with
(uStarTh <- EddyProc.C$sGetEstimatedUstarThresholdDistribution())

# display ustar scenarios
EddyProc.C$useSeaonsalUStarThresholds()
```

Plot detected ustar thresholds over time

Seasonal ustar 50% thresholds
```{r}
uStarTh %>%
  na.omit() %>%
  ggplot(aes(x = season)) +
  geom_point(aes(y = `50%`)) +
  geom_errorbar(aes(ymin = `16%`, ymax = `84%`),
                linewidth = 1, alpha = 0.2, width = 0) +
  theme_classic()
```

Annual ustar 50% thresholds
```{r}
uStarTh %>%
  filter(aggregationMode == 'year') %>%
  ggplot(aes(x = seasonYear)) +
  geom_point(aes(y = `50%`)) +
  geom_errorbar(aes(ymin = `16%`, ymax = `84%`),
                linewidth = 1, alpha = 0.2, width = 0)  +
  theme_classic()
 
```
## MDS GAP-FILLING

### NEE

**I have not dealt with the issue of negative values filled by MDS during night**
```{r}
EddyProc.C$sMDSGapFillUStarScens(fluxVar = 'NEE', 
                                 isFilterDayTime = TRUE,
                                 isFlagEntryAfterLowTurbulence= TRUE)
grep("NEE_.*_f$",names(EddyProc.C$sExportResults()), value = TRUE)
grep("NEE_.*_fsd$",names(EddyProc.C$sExportResults()), value = TRUE)
EddyProc.C$sPlotFingerprint('NEE_U16_orig')
EddyProc.C$sPlotFingerprint('NEE_U50_orig')
EddyProc.C$sPlotFingerprint('NEE_U84_orig')
EddyProc.C$sPlotFingerprint('NEE_U16_f')
EddyProc.C$sPlotFingerprint('NEE_U50_f')
EddyProc.C$sPlotFingerprint('NEE_U84_f')
```

Assessment of how many data points get filtered by Ustar filtering
```{r}
cc <- EddyProc.C$sDATA$Ustar < EddyProc.C$sTEMP$`Ustar_U50_Thres`
print(paste('Ustar threshold 50%:',sum(cc, na.rm=TRUE), 'datapoints (=',round(sum(cc,na.rm=TRUE)/length(cc)*100, digits=2),'%) are below the seasonal ustar threshold'))
# see which of the datapoints are already NAs
na <- which(is.na(EddyProc.C$sDATA$NEE))
us <- which(cc==TRUE)
matching.pos <- intersect(na,us) # these values are already NA in the data
diff <- length(us) - length(matching.pos) # subtract them from the original ustar length
print(paste('Ustar threshold 50%:',diff, 'datapoints (=',round(diff/length(cc)*100, digits=2),'%) get additionaly set to NA by ustar filtering'))
dummy <- EddyProc.C$sDATA$NEE
dummy[us] <- NA
print(paste('Data availability NEE before ustar filtering:',round(sum(!is.na(EddyProc.C$sDATA$NEE))/length(EddyProc.C$sDATA$NEE)*100, digits=2),'%'))
print(paste('Data availability NEE after ustar (U50) filtering:',round(sum(!is.na(dummy))/length(dummy)*100, digits=2),'%'))
```

### LE and H

```{r}
EddyProc.C$sMDSGapFill('LE', FillAll = TRUE, isVerbose = TRUE)
EddyProc.C$sPlotFingerprint('LE_orig')
EddyProc.C$sPlotFingerprint('LE_f')
EddyProc.C$sMDSGapFill('H', FillAll = TRUE, isVerbose = TRUE)
EddyProc.C$sPlotFingerprint('H_orig')
EddyProc.C$sPlotFingerprint('H_f')
```
## NEE PARTITIONING

```{r}
# We first need to gap-fill Tair, otherwise it could give error, although it should be without gaps
EddyProc.C$sMDSGapFill('Tair', FillAll = FALSE)
# NIGHT-TIME PARTITIONING
EddyProc.C$sMRFluxPartition(suffix = 'U16')
EddyProc.C$sMRFluxPartition(suffix = 'U84') 
EddyProc.C$sMRFluxPartition(suffix = 'U50')
EddyProc.C$sPlotFingerprint('GPP_U50_f')
```


## EXPORT RESULTS

Select the new columns to be exported
```{r}
FilledEddyData.F <- EddyProc.C$sExportResults()

# keep the timestamp information and the calculated VPD:
FilledEddyData.F$TIMESTAMP_END <- EddyData.F$TIMESTAMP
FilledEddyData.F = FilledEddyData.F %>%
  dplyr::select(TIMESTAMP_END, Ustar_U16_Thres, Ustar_U50_Thres, Ustar_U84_Thres, 
                NEE_U16_f, NEE_U50_f, NEE_U84_f, 
                GPP_U16_f, GPP_U50_f,GPP_U84_f,
                Reco_U16, Reco_U50, Reco_U84,
                LE_f, H_f)
head(FilledEddyData.F)

```

Save as csv

```{r}
# write csv file
write_csv(FilledEddyData.F, './51.1_UstarThresholds_MDS-gapfilled_NEE-partitioned.csv')
```

